---
title: "tnbc_r"
output: html_document
---


  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# setup
```{r libraries}
library(gridExtra)
library(grid)
# library(ggstance)
library(colorspace)
library(tidyverse)
library(dplyr)
library(readxl)
library(survival)
library(survminer)
library(rms)
library(knitr)
library(rmdformats)
library(xtable)
library(ggplot2)
library(reticulate)
library(timeROC)

# options(xtable.floating = FALSE, xtable.timestamp = "", xtable.comment = FALSE)
# options(max.print="75")
# knitr::opts_chunk$set(echo=FALSE, cache=FALSE, prompt=FALSE,tidy=TRUE,
#                       comment=NA,message=FALSE,warning=FALSE)
# opts_knit$set(width=75)
```

```{r import data}
data <- read.csv(file="data_tnbc.csv")

temp <- subset(data, pTNM!= 'NA', select=c(RSI, pTNM, Event_OS, Time_OS, Event, Time, GARD, n, d, Source))
temp$stage <- rep(NA, nrow(temp))
temp[temp$pTNM == "I" | temp$pTNM == "II", ][, "stage"] <- "low"
temp[temp$pTNM == "III", ][, "stage"] <- "high"
# summary(as.factor(temp$pTNM))
temp$Source[temp$Source == 'MCC_wboost'] <- "MCC"
temp$Source[temp$Source == 'NKI_wboost'] <- "NKI"
temp$GARDneg = temp$GARD*(-1)
temp$EQD2 <- temp$n*temp$d * (temp$d+10) / (2+10) # d=2,  alpha/beta = 10
temp$GARD <- temp$n*temp$d * (0.05*(temp$d-2)-log(temp$RSI)/2) #n*d=2

# OS only minus a couple of patients
temp1 <- subset(temp, stage != 'NA')

```

```{r define summary.fit}
# ---- summary of rms fit function ------ #
summary.fit <- function(fit){
  sum.tbl <- tibble("coef"=0,"exp(coef)"=0,"se"=0,"z"=0,"chi-sq"=0,"p"=0, "n"=0,"n.event"=0,"Dxy"=0)
  coef <- unname(fit$coefficients)[1]
  sum.tbl["coef"] <- coef
  sum.tbl["exp(coef)"] <- exp(coef)[1]
  se <- unname(sqrt(fit$var))[1]
  sum.tbl["se"] <- se
  zScore <- coef/se
  sum.tbl["z"] <- zScore
  sum.tbl["p"] <- (1-pnorm(abs(zScore),0,1))*2
  sum.tbl["chi-sq"] <- anova(fit)[1,1]
  sum.tbl <- round(sum.tbl,digits=3)
  sum.tbl["n"] <- unname(fit$stats[1])
  sum.tbl["n.event"] <- unname(fit$stats[2])
  sum.tbl["Dxy"] <- unname(fit$stats[9])
  sum.tbl
}

```

```{r staging}
library(Hmisc)
temp <- subset(data, pTNM!= 'NA', select=c(RSI, pTNM, Event_OS, Time_OS, Event, Time, GARD))
# temp$GARDneg = temp$GARD*(-1)

rcorr.cens(temp$GARD, Surv(temp$Time,temp$Event))
rcorr.cens(temp$pTNM, Surv(temp$Time,temp$Event))
rcorr.cens(temp$TNM_flipped, Surv(temp$Time,temp$Event))

nrow(subset(temp, pTNM==1 & GARD>=80))
histogram(~temp$GARD|temp$pTNM,layout=c(1,3),nint=20,scales=list(relation='free'),xlab="GARD",ylab="Percent of Patients within TNM Stage")

```

# Dose-GARD Distribution Comparison
```{r EQD GARD boxplots}
library(patchwork)


gardplot <- ggplot(temp) + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1), 
        axis.text.y=element_blank(),axis.ticks.y=element_blank()) + 
  xlim(0,80) +
  xlab("GARD") + ylab("") +
  geom_boxplot(aes(x=GARD, y=2), color = 'gray50', outlier.shape = NA) +
  geom_jitter(aes(x=GARD, y=2, color=Source), height = .15, size=0.8,alpha=.4)
 

eqdplot <- ggplot(temp) + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1), 
        axis.text.y=element_blank(),axis.ticks.y=element_blank()) + 
  xlim(0,80) +
  xlab("EQD2 (Gy)") + ylab("") +
  geom_boxplot(aes(x=EQD2, y=1), color = 'gray50', outlier.shape = NA) +
  geom_jitter(aes(x=EQD2, y=1, color=Source), height = .15, size=0.8, alpha=.4)
  

eqdplot / gardplot

```


# KM curves + cutpoint
```{r overall KM}

# km_os_unfilt <- 
ggsurvplot(fit=survfit(Surv(Time_OS, Event_OS) ~ 1, data=temp),
           xlab = "Time (months)", xlim = c(0, 10), break.time.by = 1,
           conf.int=FALSE, risk.table=TRUE,
           ggtheme = theme(aspect.ratio = 0.5),
           tables.theme =  theme(aspect.ratio = 0.06),
           title = "OS"
           )


# km_lc_unfilt <- 
ggsurvplot(fit=survfit(Surv(Time, Event) ~ 1, data=temp),
           xlab = "Time (months)", xlim = c(0, 10), break.time.by = 1,
           conf.int=FALSE, risk.table=TRUE,
           ggtheme = theme(aspect.ratio = 0.5),
           tables.theme =  theme(aspect.ratio = 0.06),
           title = "LC"
           )
```

```{r stratified}

os_med = round(10*median(temp1$GARD))/10
temp1$OS_strat = ""
temp1[temp1$GARD >= os_med, ][, "OS_strat"] <- "High GARD"
temp1[temp1$GARD < os_med, ][, "OS_strat"] <- "Low GARD"
temp$LC_strat = ""
lc_cut = 22
temp$LC_strat = ""
temp[temp$GARD >= lc_cut, ][, "LC_strat"] <- "High GARD"
temp[temp$GARD < lc_cut, ][, "LC_strat"] <- "Low GARD"

# OS stratified by median GARD 25.2
OS_fit <- survfit(Surv(Time_OS, Event_OS) ~ OS_strat, data=temp1)
LC_fit <- survfit(Surv(Time, Event) ~ LC_strat, data=temp)

fit <- list(LC = LC_fit, OS = OS_fit)
ggsurvplot(fit, data = temp, combine = TRUE, # Combine curves
           risk.table = TRUE,                  # Add risk table
           conf.int = FALSE,                    # Add confidence interval
           censor = FALSE,                     # Remove censor points
           tables.theme = theme_cleantable(),  # Clean risk table
           colors = c("#d95f02", "#d95f02", "1b9e77", "1b9e77"), lty=c(3,1,3,1), 
           # linetype = "strata",
           xlab = "Time (years)",
           xlim = c(0, 10),
           break.time.by = 2, # break X axis in time intervals
           legend.labs = c("LC, High GARD", "LC, Low GARD", "OS, High GARD", "OS, Low GARD")) 
    
```

``` {r separate OS and LC stratified}

ggsurvplot(OS_fit, 
           data = temp, 
           xlab = "Time (years)", xlim = c(0, 10), break.time.by = 1, 
           legend.labs = c("GARD > 25.2", "GARD < 25.2"),
           linetype = "strata",
           color="1b9e77",
           pval=TRUE, conf.int=FALSE, risk.table=TRUE) 

ggsurvplot(LC_fit, 
           data = temp, 
           xlab = "Time (years)", xlim = c(0, 10), break.time.by = 1, 
           legend.labs = c("GARD > 22", "GARD < 22"),
           linetype = "strata",
           color="#d95f02",
           pval=TRUE, conf.int=FALSE, risk.table=TRUE)


# ggsave("cut_os_unfilt.tiff", units="in", width=7, height=5, dpi=300, compression = 'lzw')


```

# AUC
```{r ROC for TNM - WILL PROBABLY DELETE}
# library(ROCit)
# ROCit_gard <- rocit(score=temp$GARD,class=temp$Event_flipped)
# plot(ROCit_gard, values=F)
# summary(ROCit_gard)

# install.packages('PRROC')
# library(PRROC)
# PRROC_obj <- roc.curve(scores.class0 = temp$GARD, weights.class0=temp$Event, curve=TRUE)
# plot(PRROC_obj)

t=10
# high_die = nrow(subset(temp, stage=='high' & Event_OS==1 & Time_OS<=t))
# high_live = nrow(subset(temp, stage=='high')) - high_die
# low_die = nrow(subset(temp, stage=='low' & Event_OS==1 & Time_OS<=t))
# low_live = nrow(subset(temp, stage=='low')) - low_die
high_die = nrow(subset(temp, stage=='high' & Event==1 & Time<=t))
high_live = nrow(subset(temp, stage=='high')) - high_die
low_die = nrow(subset(temp, stage=='low' & Event==1 & Time<=t))
low_live = nrow(subset(temp, stage=='low')) - low_die
sens_tnm = high_die/(high_die + low_die)
spec_tnm = spec_tnm = low_live/(low_live + high_live)
auc = round((sens_tnm + spec_tnm)*50,1)

myf<-function(x, x1=(1-spec_tnm), y1=sens_tnm, m=sens_tnm/(1-spec_tnm)){ 
  firstInds<-intersect(which(x >= 0), which(x < x1)) 
  secondInds<-intersect(which(x >= x1), which(x < 1))
  y<-x
  y[firstInds]<-m*x[firstInds]
  y[secondInds]<-((1-sens_tnm)*x[secondInds]+spec_tnm+sens_tnm-1)/spec_tnm
  y
}
x<-seq(0,1,0.001)
# plot(x, myf(x))
lab = paste("ROC at time t=",toString(t),", AUC=",auc)
plot(x, myf(x), ylab="Sensitivity", xlab="1-Specificity", main=lab, col="white"); 
lines(x, myf(x), col="blue");
segments(x0=0,y0=0,x1=1,y1=1, lty=2)


# t = seq(1, floor(max(temp$Time)),1)
# sens_tnm = vector()
# spec_tnm = vector()
# for (i in t) {
#    high_recur = nrow(subset(temp, stage=='high' & Event==1 & Time<=t(i)))
#    high_lc = nrow(subset(temp, stage=='high')) - high_recur
#    low_recur = nrow(subset(temp, stage=='low' & Event==1 & Time<=t(i)))
#    low_lc = nrow(subset(temp, stage=='low')) - low_recur
#    sens_tnm(i) = high_recur/(high_recur + low_recur)
#    spec_tnm(i) = low_lc/(low_lc + high_lc)
# }

```

```{r TNM nomo LC}
dd <- datadist(temp)
options(datadist='dd')
f<-cph(Surv(Time, Event) ~ pTNM,
       data = temp, surv=T, x=T, y=T)

temp$lp <- f[["linear.predictors"]]

timeROC_tnmLC<-timeROC(T=temp$Time, delta=temp$Event, marker=temp$lp,
                  cause=1,weighting="marginal",
                  times=seq(0,10,1), 
                  iid=TRUE)

plot(timeROC_tnmLC, time= 5)
plot(timeROC_tnmLC, time= 10)

timeROC_tnmLC[["AUC"]][6] # t=5
confint(timeROC_tnmLC)[["CI_AUC"]][5,] # t=5
timeROC_tnmLC[["AUC"]][11] # t=10
confint(timeROC_tnmLC)[["CI_AUC"]][10,] # t=10
```

```{r TNM nomo OS}
dd <- datadist(temp1)
options(datadist='dd')
f<-cph(Surv(Time, Event) ~ pTNM,
       data = temp1, surv=T, x=T, y=T)

temp1$lp <- f[["linear.predictors"]]

timeROC_tnmOS<-timeROC(T=temp1$Time, delta=temp1$Event, marker=temp1$lp,
                  cause=1,weighting="marginal",
                  times=seq(0,10,1), 
                  iid=TRUE)

plot(timeROC_tnmOS, time= 5)
# plot(timeROC_tnmOS, time= 10)

timeROC_tnmOS[["AUC"]][6] # t=5
confint(timeROC_tnmOS)[["CI_AUC"]][5,] # t=5
# timeROC_tnmOS[["AUC"]][11] # t=10
# confint(timeROC_tnmOS)[["CI_AUC"]][10,] # t=10
```

```{r GARD LC}
temp$Event_flipped <- rep(NA, nrow(temp))
temp[temp$Event == 0, ][, "Event_flipped"] <- 1
temp[temp$Event == 1, ][, "Event_flipped"] <- 0

timeROC_gardLC <- timeROC(T=temp$Time, delta=temp$Event, marker=temp$GARDneg,
                  cause=1,weighting="marginal",
                  times=seq(0,11,1), 
                  iid=TRUE)

plot(timeROC_gardLC, time= 5)
plot(timeROC_gardLC, time= 10)

timeROC_gardLC[["AUC"]][6] # t=5
confint(timeROC_gardLC)[["CI_AUC"]][5,] # t=5
timeROC_gardLC[["AUC"]][11] # t=10
confint(timeROC_gardLC)[["CI_AUC"]][10,] # t=10
```

```{r GARD OS}
temp1$Event_flipped <- rep(NA, nrow(temp1))
temp1[temp1$Event == 0, ][, "Event_flipped"] <- 1
temp1[temp1$Event == 1, ][, "Event_flipped"] <- 0

timeROC_gardOS <- timeROC(T=temp1$Time, delta=temp1$Event, marker=temp1$GARDneg,
                  cause=1,weighting="marginal",
                  times=seq(0,11,1), 
                  iid=TRUE)

plot(timeROC_gardOS, time=5)
# plot(timeROC_gardOS, time=10)

timeROC_gardOS[["AUC"]][6] # t=5
confint(timeROC_gardOS) [["CI_AUC"]][5,] # t=5
# timeROC_gardOS[["AUC"]][11] # t=10
# confint(timeROC_gardOS)[["CI_AUC"]][10,] # t=10
```

```{r combined LC/OS 5yr}
# need to run the earlier blocks first
time5 = 5

gard_xLC5 <- timeROC_gardLC$FP[,which(timeROC_gardLC$times==time5)]
gard_yLC5 <- timeROC_gardLC$TP[,which(timeROC_gardLC$times==time5)]
leg1 = paste("GARD, LC, AUC =",
             toString(round(timeROC_gardLC$AUC[which(timeROC_gardLC$times==time5)]*100,1)))

tnm_xLC5 <- timeROC_tnmLC$FP[,which(timeROC_tnmLC$times==time5)]
tnm_yLC5 <- timeROC_tnmLC$TP[,which(timeROC_tnmLC$times==time5)]
leg2 = paste("TNM, LC, AUC =",
             toString(round(timeROC_tnmLC$AUC[which(timeROC_tnmLC$times==time5)]*100,1)))

gard_xOS5 <- timeROC_gardOS$FP[,which(timeROC_gardOS$times==time5)]
gard_yOS5 <- timeROC_gardOS$TP[,which(timeROC_gardOS$times==time5)]
leg3 = paste("GARD, OS, AUC =",
             toString(round(timeROC_tnmOS$AUC[which(timeROC_tnmOS$times==time5)]*100,1)))

tnm_xOS5 <- timeROC_tnmOS$FP[,which(timeROC_tnmOS$times==time5)]
tnm_yOS5 <- timeROC_tnmOS$TP[,which(timeROC_tnmOS$times==time5)]
leg4 = paste("TNM, OS, AUC =",
             toString(round(timeROC_tnmOS$AUC[which(timeROC_tnmOS$times==time5)]*100,1)))

plot(0,0,type="l",xlim=c(0,1),ylim=c(0,1),xlab="1-Specificity",ylab="Sensitivity")
lines(gard_xLC5, gard_yLC5, col="#d95f02",lty=1)
lines(tnm_xLC5, tnm_yLC5, col="#d95f02",lty=3)
lines(gard_xOS5, gard_yOS5, col='#1b9e77',lty=1)
lines(tnm_xOS5, tnm_yOS5, col='#1b9e77',lty=3)
legend(.57, .3, legend=c(leg1, leg2, leg3, leg4),
       col=c("#d95f02", "#d95f02", '#1b9e77', '#1b9e77'), lty=c(3,1,3,1), cex=0.8) # "black" '#00BA38', 
abline(0,1,lty=2)
```

```{r combined LC 10yr}
# no OS bc too few at risk at 10yrs
# need to run the earlier blocks first

time10 = 10

tnm_xLC10 <- timeROC_tnmLC$FP[,which(timeROC_tnmLC$times==time10)]
tnm_yLC10 <- timeROC_tnmLC$TP[,which(timeROC_tnmLC$times==time10)]
leg1 = paste("TNM, LC, AUC =",
             toString(round(timeROC_tnmLC$AUC[which(timeROC_tnmLC$times==time10)]*100,1)))

gard_xLC10 <- timeROC_gardLC$FP[,which(timeROC_gardLC$times==time10)]
gard_yLC10 <- timeROC_gardLC$TP[,which(timeROC_gardLC$times==time10)]
leg2 = paste("GARD, LC, AUC =",
             toString(round(timeROC_gardLC$AUC[which(timeROC_gardLC$times==time10)]*100,1)))

plot(0,0,type="l",xlim=c(0,1),ylim=c(0,1),xlab="1-Specificity",ylab="Sensitivity")
lines(tnm_xLC10, tnm_yLC10, col="#d95f02",lty=3)
lines(gard_xLC10, gard_yLC10, col="#d95f02",lty=1)
legend(.57, .3, legend=c(leg1, leg2),
       col=c("#d95f02", "#d95f02"), lty=c(3,1), cex=0.8) 
abline(0,1,lty=2)
```


# cox 
```{r UVA GARD OS}
temp_filt <- temp %>%
  filter(Source == "MCC")

# discrete
midpt = median(temp_filt$GARD)
temp_filt$GARD_dichotomic <- rep(NA, nrow(temp_filt))
temp_filt[temp_filt$GARD >= midpt, ][, "GARD_dichotomic"] <- "High GARD"
temp_filt[temp_filt$GARD < midpt, ][, "GARD_dichotomic"] <- "Low GARD"
temp_filt$GARD_dichotomic = factor(temp_filt$GARD_dichotomic, levels = c("Low GARD","High GARD"))
f_discrete <- cph(Surv(Time_OS, Event_OS) ~ GARD_dichotomic, data=temp_filt, x=T, y=T, surv=T)
f_discrete

# cts
f_cont <- cph(Surv(Time_OS, Event_OS) ~ GARD, data=temp, x=T, y=T, surv=T)
f_cont
```

```{r UVA GARD LC}
temp_filt <- temp

#discrete
cut = 22
temp_filt$GARD_dichotomic <- rep(NA, nrow(temp_filt))
temp_filt[temp_filt$GARD >= cut, ][, "GARD_dichotomic"] <- "High GARD"
temp_filt[temp_filt$GARD < cut, ][, "GARD_dichotomic"] <- "Low GARD"
temp_filt$GARD_dichotomic = factor(temp_filt$GARD_dichotomic, levels = c("Low GARD","High GARD"))
f_discrete <- cph(Surv(Time_OS, Event_OS) ~ GARD_dichotomic, data=temp_filt, x=T, y=T, surv=T)
f_discrete

#cts
f_cont <- cph(Surv(Time, Event) ~ GARD, data=temp, x=T, y=T, surv=T)
f_cont
```

```{r individual LC cox}
f_LC<-cph(Surv(Time, Event) ~ GARD, data=data, x=T, y=T, surv = T)
p_LC<-Predict(f_LC, GARD = seq(1,80,by=1))
p_LC<-data.frame(p_LC) %>%
   rename(yhatLC = 'yhat') %>%
   rename(lowerLC = 'lower') %>%
   rename(upperLC = 'upper')
stats_LC <- summary.fit(f_LC) 
stats_LC <- stats_LC %>%
  rename(chisq = 'chi-sq') 

ggplot(p_LC) + ggtitle("Local Control") + geom_line(aes(x=GARD,y=yhatLC), color = '#d95f02') +
geom_abline(intercept=0, slope=0) +
geom_ribbon(aes(x=GARD,ymax=upperLC,ymin=lowerLC, fill = '#d95f02'), alpha=.2) + 
scale_x_continuous(expand = expansion(mult=0, add=0)) +
# facet_wrap(~Outcome, nrow = 1) +
xlab("GARD") + 
ylab("log Relative Hazard") +
theme_classic() +  theme(legend.position = "none") +
theme(panel.border = element_rect(fill = rgb(1,1,1,0),size = .75)) + geom_label(data = stats_LC,
             fill = rgb(.5,.5,.5,.1), hjust="left",
             size=4, label.r = unit(2,"points"),
             label.size = 0, show.legend = FALSE,
             aes(label = sprintf("Coef = %.2f \nHR = %.2f \np = %.2f", coef, exp(coef), p), x=5, y=-3))

```

```{r individual OS cox}
  
f_OS<-cph(Surv(Time_OS, Event_OS) ~ GARD, data=data, x=T, y=T, surv = T)
p_OS<-Predict(f_OS, GARD = seq(1,80,by=1))
p_OS<-data.frame(p_OS) %>%
   rename(yhatOS = yhat) %>%
   rename(lowerOS = lower) %>%
   rename(upperOS = upper)
stats_OS <- summary.fit(f_OS) 
stats_OS <- stats_OS %>%
  rename(chisq = 'chi-sq') 

# individual OS cox
plotOS <- ggplot(p_OS) + geom_line(aes(x=GARD,y=yhat), color = '#1b9e77') +
   geom_abline(intercept=0, slope=0) +
   geom_ribbon(aes(x=GARD, ymax=upper, ymin=lower, fill = '#1b9e77'), alpha=.2) +
   scale_x_continuous(expand = expansion(mult=0, add=0)) +
   # facet_wrap(~Outcome, nrow = 1) +
   xlab("GARD") + ylab("log Relative Hazard") +
   theme_classic() +  theme(legend.position = "none") +
   theme(panel.border = element_rect(fill = rgb(1,1,1,0),size = .75)) + geom_label(data = stats_OS,
                fill = rgb(.5,.5,.5,.1), hjust="left",
                size=4, label.r = unit(2,"points"),
                label.size = 0, show.legend = FALSE,
                aes(label = sprintf("OS \nCoef = %.2f \nHR = %.2f \np = %.2f", coef, exp(coef), p), x=5, y=-8))
```

```{r combined Cox (run prior cox boxes)}
p_combined <- merge(p_LC,p_OS, by="GARD")
plot_combined <- ggplot(p_combined) +
   geom_line(aes(x=GARD,y=yhatLC), color = '#d95f02') +
   geom_ribbon(aes(x=GARD,ymax=upperLC,ymin=lowerLC, fill = '#1b9e77'), alpha=.1) + 
   theme(panel.border = element_rect(fill = rgb(1,1,1,0),size = .75)) + geom_label(data = stats_LC,
                fill = rgb(.5,.5,.5,.1), hjust="left",
                size=4, label.r = unit(2,"points"),
                label.size = 0, show.legend = FALSE,
                aes(label = sprintf("LC \nCoef = %.2f \nHR = %.2f \np = %.2f", coef, exp(coef), p), color='#1b9e77', x=5, y=-8)) +
   geom_line(aes(x=GARD,y=yhatOS), color = '#1b9e77') + 
   geom_abline(intercept=0, slope=0) +
   geom_ribbon(aes(x=GARD, ymax=upperOS, ymin=lowerOS, fill = '#d95f02'), alpha=.2) +
   scale_x_continuous(expand = expansion(mult=0, add=0)) + 
   xlab("GARD") + ylab("log Relative Hazard") +
   theme_classic() +  theme(legend.position = "none") +
   theme(panel.border = element_rect(fill = rgb(1,1,1,0),size = .75)) + geom_label(data = stats_OS,
                fill = rgb(.5,.5,.5,.1), hjust="left",
                size=4, label.r = unit(2,"points"),
                label.size = 0, show.legend = FALSE,
                aes(label = sprintf("OS \nCoef = %.2f \nHR = %.2f \np = %.2f", coef, exp(coef), p), color='#d95f02', x=20, y=-8)) 

plot_combined

# ggsave('cox_combined.png', plot=plot_combined, path = NULL, width = 7, height = 5, dpi=1000)

# library(patchwork)
# 
# plotLC + plotOS
```


# nomogram
```{r setup}

# temp_filt <- read.csv(file="GARD_HPVpos6.csv") %>% filter(RT == "definitive") %>% 
#   rename(AJCC8 = 'TNM8') %>%
#   rename(ECOG = 'ECOG.PS') %>% 
#   mutate(ECOG = as.character(ECOG)) %>%
#   mutate(ECOG = if_else(ECOG=='0','0','1or2')) %>%
#   mutate(smoking = ifelse(smoking==0,NA,smoking))  %>%
#   mutate(smoking = ifelse(smoking=='Never',0,1)) 
  # %>% mutate(TNM = if_else(TNM=="I","I","II/III"))
  # mutate(Time_OS = Time_OS/12) %>%

# hist(x=temp$GARD,breaks=30)

```


creating the cph model:
```{r}
dd <- datadist(temp)
options(datadist='dd')

f_OS<-cph(Surv(Time_OS, Event_OS) ~ GARD + pTNM, 
       data = temp1, surv=T, x=T, y=T)
f_OS

f_LC<-cph(Surv(Time, Event) ~ GARD + pTNM, 
       data = temp, surv=T, x=T, y=T)
f_LC
```

creating the nomogram plot
```{r}
surv_LC <- Survival(f_LC)
surv_OS <- Survival(f_OS)
surv2 <- function(x) surv_LC(2, lp=x)
surv5 <- function(x) surv_LC(5, lp=x)
surv10 <- function(x) surv_LC(10, lp=x)

# surv percentages to be included on plot (0.6 = 60% surv)
ss<-c(0.2,0.3, 0.4,0.5, 0.6,0.7,0.8,0.9,0.95,0.98)

# can include multiple functions in the list, but will need to update other arguments to match the number of inputs
nom <- nomogram(f, 
                fun = list(surv5), 
                funlabel = c('5 yr surv'), 
                fun.at = list(ss) )

# plot(nom, xfrac = 0.4)

dd <- datadist(temp_filt)
options(datadist='dd')
f_OS<-cph(Surv(Time_OS, Event_OS) ~ GARD + pTNM,
       data = temp, surv=T, x=T, y=T)
f_LC<-cph(Surv(Time, Event) ~ GARD + pTNM,
       data = temp, surv=T, x=T, y=T)

temp$lp_OS <- f_OS[["linear.predictors"]]
temp$lp_LC <- f_LC[["linear.predictors"]]

timeROC_nomo_os<-timeROC(T=temp$Time_OS, delta=temp$Event_OS, marker=temp$lp_OS,
                  cause=1,weighting="marginal",
                  times=seq(0,10,.1), 
                  iid=TRUE)
timeROC_nomo_lc<-timeROC(T=temp$Time, delta=temp$Event, marker=temp$lp_LC,
                  cause=1,weighting="marginal",
                  times=seq(0,10,.1),
                  iid=TRUE)

# plot(timeROC_nomo, time= 2)
plot(timeROC_nomo_lc, time= 10)
ggsave("timeROC_nomo.png", units="in", width=7, height=5, dpi=300, compression = 'lzw')
# plot(timeROC_nomo_no, time= 5)

```

# tertiles
```{r load_data}

#not really sure if this is relevant bc Jess's version uses it to make the GARD column
surv <- Surv(time = temp$Time, event = temp$Event) 
surv_os <- Surv(time = temp1$Time_OS, event = temp1$Event_OS) 
```
## Find optimal cutpoints
Next, we'll find sweep all cutpoint options to find the optimal cutpoint for three (`sweep_cutpoints2`) groups. 
```{r load_results}
sweep_cutpoints2 <- function(tidy_data, surv_obj){
   cutpoints_first = quantile(tidy_data$GARD, prob = seq(0, 1, 0.01))[-(1:5)]
   cutpoints_first = cutpoints_first[-(96:101)]
   cutpoint_first_vec <- c()
   cutpoint_second_vec <- c()
   low_n_vec <- c()
   middle_n_vec <- c()
   high_n_vec <- c()
   chisq_vec <- c()
   for (i in 1:length(cutpoints_first)){
      GARD_after_cutoff <- tidy_data$GARD[tidy_data$GARD > min(cutpoints_first)]
      cutpoints_second <- quantile(GARD_after_cutoff, prob = seq(0, 1, 0.01))[-(1:5)]
      cutpoints_second <- cutpoints_second[-(96:101)]
      for (j in 1:length(cutpoints_second)){
         tidy_data_labs <- tidy_data %>%
            mutate(pred_group = case_when(GARD <= cutpoints_first[i] ~ "low", 
                                          GARD > cutpoints_first[i] &
                                             GARD <= cutpoints_second[j] ~ "middle", 
                                          GARD > cutpoints_second[j] ~ "high"))
         survdiff_results <- survdiff(surv_obj ~ pred_group, data = tidy_data_labs)
         chisq_vec <- c(chisq_vec, survdiff_results$chisq)
         cutpoint_first_vec <- c(cutpoint_first_vec, cutpoints_first[[i]])
         cutpoint_second_vec <- c(cutpoint_second_vec, cutpoints_second[[j]])
         low_n_vec <- c(low_n_vec, nrow(tidy_data_labs %>% filter(pred_group == "low")))
         middle_n_vec <- c(middle_n_vec, nrow(tidy_data_labs %>% filter(pred_group == "middle")))
         high_n_vec <- c(high_n_vec, nrow(tidy_data_labs %>% filter(pred_group == "high")))
      }
   }
   data.frame(cutpoint1 = cutpoint_first_vec,
              cutpoint2 = cutpoint_second_vec,
              low_n = low_n_vec,
              middle_n = middle_n_vec,
              high_n = high_n_vec,
              chisq = chisq_vec) %>%
      rowwise() %>%
      mutate(min_n = min(low_n, middle_n, high_n)) %>%
      ungroup()
}

cutpoints <- sweep_cutpoints2(temp, surv_os)
cutpoints2 <- sweep_cutpoints2(temp2, surv)

cutpoints_filtered <- cutpoints %>%
   filter(min_n >= 4) %>% #require at least 4 samples in each group
   filter(chisq > 10) # == max(chisq)) #   #cutpoints with the highest chisq result

cutpoints_filtered2 <- cutpoints2 %>%
   filter(min_n >= 4) %>% #require at least 4 samples in each group
   filter(chisq > 10) # == max(chisq) #cutpoints with the highest chisq result

```


## Classify predictions
Now, we'll classify each of our samples as high, medium, or low. 
```{r classify predictions}

# LC
cutpoints_filtered2 <- cutpoints_filtered2[order(-cutpoints_filtered2$chisq),]
rank <- 1
double_cutpoint_best <- c(cut1 = cutpoints_filtered2[rank, ]$cutpoint1, 
                          cut2 = cutpoints_filtered2[rank, ]$cutpoint2)

lab1 = paste("GARD <",toString(trunc(10*cutpoints_filtered2[1, ]$cutpoint1)/10))
lab2 = paste("GARD ",toString(trunc(10*cutpoints_filtered2[1,]$cutpoint1)/10),"-",toString(trunc(10*cutpoints_filtered2[1, ]$cutpoint2)/10))
lab3 = paste("GARD >",toString(trunc(10*cutpoints_filtered2[1, ]$cutpoint2)/10))

temp2 <- temp2 %>%
      mutate(
          pred_3grp_tert2 = factor(case_when(GARD < quantile(GARD, 0.33) ~ "1_Low",
                                GARD >= quantile(GARD, 0.33) & GARD < quantile(GARD, 0.66) ~ "2_Middle",
                                GARD >= quantile(GARD, 0.66) ~ "3_High"),
                                ordered = TRUE),
          pred_3grp_optimal2 = case_when(GARD <= double_cutpoint_best["cut1"] ~ "1_Low",
                                GARD > double_cutpoint_best["cut1"] &
                                   GARD <= double_cutpoint_best["cut2"] ~ "2_Middle",
                                GARD > double_cutpoint_best["cut2"] ~ "3_High"),
          pred_3grp_25502 = factor(case_when(GARD < quantile(GARD, 0.25) ~ "1_Low",
                                GARD >= quantile(GARD, 0.25) &
                                   GARD < quantile(GARD, 0.75) ~ "2_Middle",
                                GARD >= quantile(GARD, 0.75) ~ "3_High"), 
                                ordered = TRUE))


# OS
df <- cutpoints_filtered[order(-cutpoints_filtered$chisq),]

# they all cut the data with the same samples in each group
# taking the first value to get a single value for each cutpoint
rank <- 1
double_cutpoint_best <- c(cut1 = df[rank, ]$cutpoint1, 
                          cut2 = df[rank, ]$cutpoint2)

lab1 = paste("GARD <",toString(trunc(10*df[rank, ]$cutpoint1)/10))
lab2 = paste("GARD ",toString(trunc(10*df[rank, ]$cutpoint1)/10),"-",toString(trunc(10*df[rank, ]$cutpoint2)/10))
lab3 = paste("GARD >",toString(trunc(10*df[rank, ]$cutpoint2)/10))
temp <- temp %>%
      mutate(
          pred_3grp_tert = factor(case_when(GARD < quantile(GARD, 0.33) ~ "1_Low",
                                GARD >= quantile(GARD, 0.33) & GARD < quantile(GARD, 0.66) ~ "2_Middle",
                                GARD >= quantile(GARD, 0.66) ~ "3_High"),
                                ordered = TRUE),
          pred_3grp_optimal = case_when(GARD <= double_cutpoint_best["cut1"] ~ "1_Low",
                                GARD > double_cutpoint_best["cut1"] &
                                   GARD <= double_cutpoint_best["cut2"] ~ "2_Middle",
                                GARD > double_cutpoint_best["cut2"] ~ "3_High"),
          pred_3grp_2550 = factor(case_when(GARD < quantile(GARD, 0.25) ~ "1_Low",
                                GARD >= quantile(GARD, 0.25) &
                                   GARD < quantile(GARD, 0.75) ~ "2_Middle",
                                GARD >= quantile(GARD, 0.75) ~ "3_High"), 
                                ordered = TRUE))
```


## Plot Results
```{r Optimal 3 group cutpoints}
# OS
km_tert_opt <- survfit(surv_os ~ temp$pred_3grp_optimal)
gg_grp3_opt <- ggsurvplot(km_tert_opt, data = temp, size = 1,
   conf.int = FALSE, # Add confidence interval
   pval = TRUE, # Add p-value
   risk.table = TRUE, # Add risk table
   risk.table.height = 0.25, # change if you have >2 groups
   risk.table.y.text.col = T, # show colors in table listing
   ncensor.plot = FALSE,
   ncensor.plot.height = 0.25,
   xlab = "Time in years", # adjust X axis label
   xlim = c(0, 18),
   break.time.by = 1, # break X axis in time intervals
   legend.labs = c(lab1, lab2, lab3), #c("Low GARD", "Middle GARD", "High GARD"),
   title = "TNBC OS grouped by GARD tertile (optimal)",
   ggtheme = theme_bw() + theme(plot.title = element_text(hjust = 0.5))
   )
gg_grp3_opt


# LC
km_tert_opt2 <- survfit(surv ~ temp$pred_3grp_optimal)
gg_grp3_opt2 <- ggsurvplot(km_tert_opt2, data = temp, size = 1,
   conf.int = FALSE, # Add confidence interval
   pval = TRUE, # Add p-value
   risk.table = TRUE, # Add risk table
   risk.table.height = 0.25, # change if you have >2 groups
   risk.table.y.text.col = T, # show colors in table listing
   ncensor.plot = FALSE,
   ncensor.plot.height = 0.25,
   xlab = "Time in years", # adjust X axis label
   xlim = c(0, 18),
   break.time.by = 1, # break X axis in time intervals
   legend.labs = c(lab1, lab2, lab3), #c("Low GARD", "Middle GARD", "High GARD"),
   title = "TNBC LC grouped by GARD tertile (optimal)",
   ggtheme = theme_bw() + theme(plot.title = element_text(hjust = 0.5))
   )
gg_grp3_opt2
```

```{r Tertiles by 33%}
km_tert_by33 <- survfit(surv_os ~ temp$pred_3grp_tert)
gg_grp3_tert <- ggsurvplot(km_tert_by33, data = temp, size = 1,
   conf.int = FALSE, # Add confidence interval
   pval = TRUE, # Add p-value
   risk.table = TRUE, # Add risk table
   risk.table.height = 0.25, # change if you have >2 groups
   risk.table.y.text.col = T, # show colors in table listing
   ncensor.plot = FALSE,
   ncensor.plot.height = 0.25,
   xlab = "Time in years", # adjust X axis label
   xlim = c(0, 18),
   break.time.by = 1, # break X axis in time intervals
   legend.labs = c("Low GARD", "Middle GARD", "High GARD"),
   title = "TNBC OS grouped by GARD tertile 33/33/33",
   ggtheme = theme_bw() + theme(plot.title = element_text(hjust = 0.5))
   )
gg_grp3_tert


km_tert_by33_2 <- survfit(surv ~ temp$pred_3grp_tert)
gg_grp3_tert2 <- ggsurvplot(km_tert_by33_2, data = temp, size = 1,
   conf.int = FALSE, # Add confidence interval
   pval = TRUE, # Add p-value
   risk.table = TRUE, # Add risk table
   risk.table.height = 0.25, # change if you have >2 groups
   risk.table.y.text.col = T, # show colors in table listing
   ncensor.plot = FALSE,
   ncensor.plot.height = 0.25,
   xlab = "Time in years", # adjust X axis label
   xlim = c(0, 18),
   break.time.by = 1, # break X axis in time intervals
   legend.labs = c("Low GARD", "Middle GARD", "High GARD"),
   title = "TNBC LC grouped by GARD tertile 33/33/33",
   ggtheme = theme_bw() + theme(plot.title = element_text(hjust = 0.5))
   )
gg_grp3_tert2
```

```{r Tertiles by 25/50/25}
km_tert_by25_50_25 <- survfit(surv_os ~ temp$pred_3grp_2550)
gg_grp3_2550 <- ggsurvplot(km_tert_by25_50_25, data = temp, size = 1,
   conf.int = FALSE, # Add confidence interval
   pval = TRUE, # Add p-value
   risk.table = TRUE, # Add risk table
   risk.table.height = 0.25, # change if you have >2 groups
   risk.table.y.text.col = T, # show colors in table listing
   ncensor.plot = FALSE,
   ncensor.plot.height = 0.25,
   xlab = "Time in years", # adjust X axis label
   xlim = c(0, 18),
   break.time.by = 1, # break X axis in time intervals
   legend.labs = c(lab1, lab2, lab3), #c("Low GARD", "Middle GARD", "High GARD"),
   title = "TNBC OS grouped by GARD, 25/50/25",
   ggtheme = theme_bw() + theme(plot.title = element_text(hjust = 0.5))
   )
gg_grp3_2550

km_tert_by25_50_25_2 <- survfit(surv ~ temp$pred_3grp_2550)
gg_grp3_2550_2 <- ggsurvplot(km_tert_by25_50_25_2, data = temp, size = 1,
   conf.int = FALSE, # Add confidence interval
   pval = TRUE, # Add p-value
   risk.table = TRUE, # Add risk table
   risk.table.height = 0.25, # change if you have >2 groups
   risk.table.y.text.col = T, # show colors in table listing
   ncensor.plot = FALSE,
   ncensor.plot.height = 0.25,
   xlab = "Time in years", # adjust X axis label
   xlim = c(0, 18),
   break.time.by = 1, # break X axis in time intervals
   legend.labs = c("Low GARD", "Middle GARD", "High GARD"),
   title = "TNBC LC grouped by GARD, 25/50/25",
   ggtheme = theme_bw() + theme(plot.title = element_text(hjust = 0.5))
   )
gg_grp3_2550_2
```
